---
title: "Power calculations for Experiment"
author: "Colette Salemi and Mark Marvin Kadigo"
date: "2023-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

##Overview

This script uses code provided by JPAL to perform retrospective power calculations. Based on sample sizes and other parameters (p-value, power, etc.), the code determines minimum detectible size.

Throughout the code focuses on the power of the coefficient of interest, sameOcc*outGroup. Additional power calculations can be run for other coefficients, such as the coefficient for sameOcc, but then the variable of interest (the interaction) should be included as a covariate. 

## Data prep

This is the made-up data Colette used to build the code. Mark, please delete the fake data production chunk immediately below and open the survey data instead. Make sure the survey data is called df.

```{r}

#Make sure dplyr is installed
#install.packages("dplyr")   #Mark making sure.
library(dplyr)

#num_rows = 3000

#df <- data.frame(
#  id = 1:num_rows,
#  exp_outgroup = sample(0:1, num_rows, replace = TRUE),  
#  exp_same_occup = sample(0:1, num_rows, replace = TRUE),
#  refugee = sample(0:1, num_rows, replace = TRUE),
#  primary = sample(0:1, num_rows, replace = TRUE),
#  male = sample(0:1, num_rows, replace = TRUE),
#  ethiopia = sample(0:1, num_rows, replace=TRUE),
#  urban = sample(0:1, num_rows, replace=TRUE),
#  contact = sample(0:1, num_rows, replace=TRUE),
#  hhsize = sample(1:7, num_rows, replace=TRUE),
#  educ_primary = sample(0:1, num_rows, replace=TRUE),
#  employed = sample(0:1, num_rows, replace=TRUE),
#  age = sample(15:64, num_rows, replace = TRUE)
#)

#Create outcome variables that are statistically different from each other.
#df <- df %>%
#  mutate(prej_aind_byR = case_when(
#    exp_outgroup == 0 & exp_same_occup == 0 ~ rnorm(num_rows, mean #= 3, sd = 0.5),
#    exp_outgroup == 1 & exp_same_occup == 0 ~ rnorm(num_rows, mean #= 15, sd = 0.5),
#    exp_outgroup == 0 & exp_same_occup == 1 ~ rnorm(num_rows, mean #= 57, sd = 0.5),
#    exp_outgroup == 1 & exp_same_occup == 1 ~ rnorm(num_rows, mean #= 103, sd = 0.5)
#  ))

#df$prej_aind_byL = df$prej_aind_byR

#head(df)
#colnames(df)

```
Create a treatment variable, an interaction between exp_same_occ and exp_outgroup

```{r}

#df$Treatment = df$exp_outgroup*df$exp_same_occ


```

##Setup

This section opens libraries and defines parameters used throughout.

First, libraries required to run the JPAL tools. pwrcalc needs to be installed from Github. Make sure all packages are installed first.
```{r}
#Only run these installations once, turn them "on" and "off" using a # character at the start of the line

#install.packages("devtools")
#devtools::install_github('vikjam/pwrcalc')  #Install this from github

#install.packages("haven")
#install.packages("ICC")
#install.packages("randomizr")
#install.packages("multiwayvcov")
#install.packages("lmtest")
#install.packages("magrittr")
#install.packages("pwrcalc")
#install.packages("Hmisc")
#install.packages("gsubfn")
#install.packages("writexl")
#install.packages("readxl")  


#Open package libraries
library(devtools)
library(haven)
library(ICC)
library(randomizr)
library(multiwayvcov)
library(lmtest)
library(magrittr)
library(pwrcalc)
library(Hmisc)
library(gsubfn)
library(writexl)
library(readxl)
#library(xlsx)
```
Opening the actual data and name it df 

```{r}
# Detect system username dynamically
detect_username <- function() {
  if (Sys.info()["sysname"] == "Windows") {
    return(Sys.getenv("USERNAME"))
  } else {
    return(Sys.getenv("USER"))
  }
}

username <- detect_username()

# Construct base path dynamically
base_directory <- file.path("C:", "Users", username)

data_subpath <- "ETH_UGA Experiment/04. Experimental Approach FAFO/09. Replication Package JDE/01. Datasets/03. Final/04_UGA_ETH_Prepared.dta"

output_subpath <- "ETH_UGA Experiment/04. Experimental Approach FAFO/09. Replication Package JDE/02. Output/Tables"

# Identify the correct cloud storage folder
possible_folders <- c("ownCloud - julie.bousquet@kuleuven.be@owncloud.gwdg.de", "Dropbox")
found_folder <- NULL

for (folder in possible_folders) {
  full_path <- file.path(base_directory, folder)
  if (dir.exists(full_path)) {
    found_folder <- full_path
    break
  }
}

if (!is.null(found_folder)) {
  file_path <- file.path(found_folder, data_subpath)
  output_folder <- file.path(found_folder, output_subpath)
} else {
  # Fall back to a simpler file path structure
  file_path <- file.path(base_directory, "Documents", "data_file.dta")
  output_folder <- file.path(base_directory, "Documents", "Tables")
}

# Ensure output directory exists
if (!dir.exists(output_folder)) {
  dir.create(output_folder, recursive = TRUE)
}

df <- read_dta(file_path)


```
Specifying the interaction treatment

```{r}
df$Treatment = df$exp_outgroup*df$exp_same_occup
```

Specify power level, significance level, ratio of treatment to control. The ratio is based on 1/4 of the sample being treatment and 3/4 nontreatment (1/3)

```{r}
power = 0.8       #SPECIFY - desired power
nratio = 0.33     #SPECIFY - the ratio of the size of the treatment group to control group
alpha =0.05       #SPECIFY - significance level
p = nratio/(1+nratio)
```

Function to detect MDS with covariates in estimation. This comes directly from JPAL. The function determines MDS.

```{r}

mde_cov <- function (dataset, outcome, covariates, treatment, N){
  t_power = qt(power, df=N-2)
  t_alpha = qt(1-alpha/2, df=N-2)
  
  cov_list <- paste(covariates, collapse = " + ")
  formula <- as.formula(paste("outcome ~ ",cov_list,sep = ""))
  fit <- lm(formula, data = dataset)
  summary(fit)
  
  res_baseline_sd <- sd(summary(fit)$residuals, na.rm=TRUE)
  res_baseline_sd                                                               
  
  mde_res <- (t_power + t_alpha) * sqrt(1 /(p*(1-p))) * sqrt(1 / N) * res_baseline_sd
  mde_res = round(mde_res, digits=2)
  
}

```

Dataframe to store results. I wrote the code so that it builds a dataframe that tells us which regression and which coefficient the power calculation is for.

```{r}

results <- data.frame(
  Group = character(0),
  Type = character(0),
  Location = character(0),
  Variable = character(0),
  MDS = numeric(0)
)
```

Specify additional variables and parameters consistent throughout
```{r}

#SPECIFY covariates in regression
cov= c("exp_outgroup", "exp_same_occup", "ethiopia", "urban", "male", "age", "hhsize", "educ_primary", "employed") 

#Data subsets
hosts = df[which(df$refugee==0),] 
refugees = df[which(df$refugee==1),]

```

## Main results 

###Results for refugees

Overall
```{r}

#Parameters to manually change for the results table
group = "Refugees" #group is either host or refugee
type  = "Main"  #type is either main regression or a particular heterogeneity analysis
location = "All" #either all locations or specific locations
variable = "Interaction" #Either the interaction term, outGroup, or sameOcc

dataset = refugees                  #SPECIFY - the dataset
outcome = dataset$prej_aind_byR     #SPECIFY - the outcome variable
treatment = dataset$Treatment       #SPECIFY - treatment variable

N = nrow(dataset) 

#Execute the function
mde_res <- mde_cov(dataset, outcome, cov, treatment, N)   

#Store results in a table
res = unlist(c(group, type, location, variable, mde_res))

res_matrix <- matrix(res, nrow = 1, byrow = TRUE,
                     dimnames = list(NULL, c("Group", "Type", "Location", "Variable", "MDS")))

# Bind the matrix as a row to the empty dataframe
results = rbind(results, as.data.frame(res_matrix))


```

Jijiga
```{r}

#Parameters to manually change for the results table
group = "Refugees" #group is either host or refugee
type  = "Main"  #type is either main regression or a particular heterogeneity analysis
location = "Jijiga" #either all locations or specific locations
variable = "Interaction" #Either the interaction term, outGroup, or sameOcc

dataset = refugees[which(refugees$region_short==2),]
#dataset = refugees[which(refugees$urban==0 & refugees$ethiopia==1),]                  
outcome = dataset$prej_aind_byL       #SPECIFY - the outcome variable
treatment = dataset$Treatment       #SPECIFY - treatment variable

N = nrow(dataset) 

#Execute the function
mde_res <- mde_cov(dataset, outcome, cov, treatment, N)   

#Store results in a table
res = unlist(c(group, type, location, variable, mde_res))

res_matrix <- matrix(res, nrow = 1, byrow = TRUE,
                     dimnames = list(NULL, c("Group", "Type", "Location", "Variable", "MDS")))

# Bind the matrix as a row to the empty dataframe
results = rbind(results, as.data.frame(res_matrix))

```

Addis Ababa
```{r}

#Parameters to manually change for the results table
group = "Refugees" #group is either host or refugee
type  = "Main"  #type is either main regression or a particular heterogeneity analysis
location = "Addis Ababa" #either all locations or specific locations
variable = "Interaction" #Either the interaction term, outGroup, or sameOcc

dataset = refugees[which(refugees$region_short==1),]
#dataset = refugees[which(refugees$urban==1 & refugees$ethiopia==1),]                  
outcome = dataset$prej_aind_byL     #SPECIFY - the outcome variable
treatment = dataset$Treatment       #SPECIFY - treatment variable

N = nrow(dataset) 

#Execute the function
mde_res <- mde_cov(dataset, outcome, cov, treatment, N)   

#Store results in a table
res = unlist(c(group, type, location, variable, mde_res))

res_matrix <- matrix(res, nrow = 1, byrow = TRUE,
                     dimnames = list(NULL, c("Group", "Type", "Location", "Variable", "MDS")))

# Bind the matrix as a row to the empty dataframe
results = rbind(results, as.data.frame(res_matrix))

```

Isingiro
```{r}

#Parameters to manually change for the results table
group = "Refugees" #group is either host or refugee
type  = "Main"  #type is either main regression or a particular heterogeneity analysis
location = "Isingiro" #either all locations or specific locations
variable = "Interaction" #Either the interaction term, outGroup, or sameOcc

dataset = refugees[which(refugees$region_short==4),]
#dataset = refugees[which(refugees$urban==0 & refugees$ethiopia==0),]                  
outcome = dataset$prej_aind_byL           #SPECIFY - the outcome variable
treatment = dataset$Treatment       #SPECIFY - treatment variable

N = nrow(dataset) 

#Execute the function
mde_res <- mde_cov(dataset, outcome, cov, treatment, N)   

#Store results in a table
res = unlist(c(group, type, location, variable, mde_res))

res_matrix <- matrix(res, nrow = 1, byrow = TRUE,
                     dimnames = list(NULL, c("Group", "Type", "Location", "Variable", "MDS")))

# Bind the matrix as a row to the empty dataframe
results = rbind(results, as.data.frame(res_matrix))

```

Kampala
```{r}

#Parameters to manually change for the results table
group = "Refugees" #group is either host or refugee
type  = "Main"  #type is either main regression or a particular heterogeneity analysis
location = "Kampala" #either all locations or specific locations
variable = "Interaction" #Either the interaction term, outGroup, or sameOcc

dataset = refugees[which(refugees$region_short==3),]
#dataset = refugees[which(refugees$urban==1 & refugees$ethiopia==0),]                   
outcome = dataset$prej_aind_byL           #SPECIFY - the outcome variable
treatment = dataset$Treatment       #SPECIFY - treatment variable

N = nrow(dataset) 

#Execute the function
mde_res <- mde_cov(dataset, outcome, cov, treatment, N)   

#Store results in a table
res = unlist(c(group, type, location, variable, mde_res))

res_matrix <- matrix(res, nrow = 1, byrow = TRUE,
                     dimnames = list(NULL, c("Group", "Type", "Location", "Variable", "MDS")))

# Bind the matrix as a row to the empty dataframe
results = rbind(results, as.data.frame(res_matrix))

```

## Exporting and Appending results to Excel

```{r}
write_xlsx(results, file.path(output_folder, "T24_Retrospective_PC.xlsx"))

# Load Excel file to create LaTeX table
excel_file <- file.path(output_folder, "T24_Retrospective_PC.xlsx")
data <- read_excel(excel_file)

# Ensure correct column names and remove unnecessary columns
colnames(data) <- c("Group", "Type", "Location", "Variable", "MDS")
data <- data %>% select(Group, Location, MDS)

# Modify the Group column to merge rows in LaTeX
data$Group[duplicated(data$Group)] <- ""

# Generate a raw LaTeX tabular format without repeating "Refugees" in every row
latex_table <- paste(
  "\\begin{tabular}{c|c|c}",
  "Group & Location & Minimum Detectable Size (MDS) \\\\",
  "\\hline \\hline",
  paste0(data$Group, " & ", data$Location, " & ", data$MDS, " \\\\", collapse = "\n"),
  "\\hline",
  "\\end{tabular}",
  sep = "\n"
)

# Save LaTeX table to a .tex file in the output folder
latex_file <- file.path(output_folder, "TA24_Retrospective_PC.tex")
write(latex_table, file = latex_file)

file.remove(file.path(output_folder, "T24_Retrospective_PC.xlsx"))

``````







